Python as a machine learning language: experiences from the \textit{scikit-learn} project

### 2013-06-05 paper.tex
Abstract + Introduction

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage[round]{natbib}

% working title, please change
\title{Python as a machine learning language:
       experiences from the scikit-learn project}

@@

\institute{Graduate School of System Informatics, Kobe University \and
           Informatics Institute, University of Amsterdam \and
           Dept. of EE \& CS, University of Liège \and
           Astronomy Department, University of Washington}

@@

\begin{abstract}
The scikit-learn project aims to create an open source
machine learning library for the Python programming language,
offering performant implementations of well-established learning algorithms,
and tying in with the ecosystem of scientific software
that has evolved around the NumPy and SciPy packages.
Since the start of the project in 2009,
it has gathered a sizeable user base.

This paper discusses design choices made
in both the programmers' interface (API) and the implementation
of scikit-learn.
It also highlights obstacles faced by users and implementers
of machine learning algorithms in the Python language.
\end{abstract}

\section{Project vision}

\subsection{General principles}

% XXX the following is more of a general introduction than a project vision
scikit-learn \citep{pedregosa2011} is an open source machine learning library
for the Python programming language,
intended to tie in with the set of numeric and scientific packages
centered around the NumPy and SciPy libraries
\Citep{vanderwalt2011, varoquaux2013scipy}.
It is a library, rather than a novel (domain-specific) programming language,
intended to be called from Python programs;
at present, it offers no command line interface or GUI.

The scikit-learn project started in 2009 and has since been developed
by a team largely consisting of scientists
(both from computer science and other fields).
Its popularity can be gauged from various sources:
% TODO: list volume on the mailing list, number of contributors,
% number of issues opened and closed, number of forks, number of surveys filled in.
at the time of writing,
the popular programmer's Q~\&{}A site StackOverflow\footnote{
  \url{http://stackoverflow.com/questions/tagged/scikit-learn}}
% update the following figure when paper is finalized
lists 322 scikit-learn-related questions.
The version control logs for the development version of scikit-learn
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 contributors to the project's source code and documentation
(after manual deduplication).

The main goal of the scikit-learn project
is to provide high-quality implementations
of ``classic'' machine learning algorithms
with a consistent API.

@@

In the remainder of the paper, we use the following notation. Training data are
denoted by \Xtr ~and test data (i.e., unseen data to which the learning
algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
$i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
such as classification or regression, we denote the training and test labels by
\ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
sample's label value).

>>>>

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage[round]{natbib}
\usepackage{xcolor}

% working title, please change
\title{Python as a machine learning language:
       experiences from the \textit{scikit-learn} project}

@@

\institute{Graduate School of System Informatics, Kobe University \and
           Informatics Institute, University of Amsterdam \and
           Department of EE \& CS, University of Liège \and
           Astronomy Department, University of Washington}

@@

\begin{abstract}
The \textit{scikit-learn} project is an increasingly popular open source machine
learning library designed to be simple and efficient, accessible to non-experts,
and reusable in various contexts. In this paper, we present and discuss our
design choices for the application programming interface (API) of the project.
In particular, we describe the \textit{estimator} interface shared by all
learning and processing units in the library and then discuss its advantages in
terms of composition and reusability. The paper also comments on implementation
details specific to the Python ecosystem and highlights obstacles faced by users
and developers of machine learning algorithms.
\end{abstract}

\section{The \textit{scikit-learn} project}

The \textit{scikit-learn} project~\citep{pedregosa2011} is an open source
machine learning library for the Python programming language. The ambition of
the project is to provide  effective and long-established machine learning tools
within a programming environment that is easy-to-use, efficient, accessible to
non-machine learning experts and reusable in various scientific areas. In
itself, the project is a library of machine learning tools including, among
others, learning algorithms, model selection tools and preprocessing procedures
rather than a novel domain-specific programming language. From its core, the
library has been designed to tie in with the set of numeric and scientific
packages centered around the Numpy and SciPy~\citep{vanderwalt2011,
varoquaux2013scipy} libraries. As such, \textit {scikit-learn} is intended to be
used within Python scripts and requires some Python programming knowledge. At
the time, it offers no command line or graphical user interface.

Started in 2009, \textit{scikit-learn} counts a large team of volunteer
international developers mainly consisting of scientists and researchers from
various areas (e.g., machine learning, natural language processing or
neuroscience). The project also greatly benefits from many occasional
contributors proposing small code and/or documentation fixes or improvements.
Because of this large base of developers, important efforts are demanded to make
the project maintainable. In particular, we emphasize on high-quality and well-documented
code  in order to ease future developments by external contributors.

Within the past few years, the project has proven to be valuable to an  ever-
increasing user base. Its popularity can be gauged from various indicators,
including hundreds of citations in scientific papers, successes in various
machine learning challenges, or statistics from our repositories and mailing
lists.

\textcolor{red}{The rest of this text is organized as follows: in Section 2, ...}

% % XXX the following is more of a general introduction than a project vision
% scikit-learn \citep{pedregosa2011} is an open source machine learning library
% for the Python programming language,
% intended to tie in with the set of numeric and scientific packages
% centered around the NumPy and SciPy libraries
% \Citep{vanderwalt2011, varoquaux2013scipy}.
% It is a library, rather than a novel (domain-specific) programming language,
% intended to be called from Python programs;
% at present, it offers no command line interface or GUI.

% The scikit-learn project started in 2009 and has since been developed
% by a team largely consisting of scientists
% (both from computer science and other fields).
% Its popularity can be gauged from various sources:
% % TODO: list volume on the mailing list, number of contributors,
% % number of issues opened and closed, number of forks, number of surveys filled in.
% at the time of writing,
% the popular programmer's Q~\&{}A site StackOverflow\footnote{
%   \url{http://stackoverflow.com/questions/tagged/scikit-learn}}
% % update the following figure when paper is finalized
% lists 322 scikit-learn-related questions.
% The version control logs for the development version of scikit-learn
% % ddaa494c116e3c16bf032003c5cccbed851733d2
% show 183 contributors to the project's source code and documentation
% (after manual deduplication).

% The main goal of the scikit-learn project
% is to provide high-quality implementations
% of ``classic'' machine learning algorithms
% with a consistent API.

@@

% In the remainder of the paper, we use the following notation. Training data are
% denoted by \Xtr ~and test data (i.e., unseen data to which the learning
% algorithm needs to be able to generalize) by \Xte.  Following the discussion above,
% we represent \Xtr ~and \Xte ~by 2-dimensional NumPy arrays or SciPy sparse
% matrices. Using NumPy/SciPy's notation, \Xtr\texttt{[i, j]} denotes the
% $i^\textrm{th}$ training sample's $j^\textrm{th}$ feature. For supervised learning tasks,
% such as classification or regression, we denote the training and test labels by
% \ytr ~and \yte, respectively. We typically store \ytr ~and \yte ~as
% 1-dimensional NumPy arrays (\ytr\texttt{[i]} denotes the $i^\textrm{th}$ training
% sample's label value).
